Supplemental figures for RLSmith Transl Psych 2025

### 2025-07-02 scripts/07.figures_markdown/supplement_v2.Rmd
Update scripts with response to reviewer comments

---
title: "Supplemental figures for RLSmith Transl Psych 2024"
author: "Rachel L Smith"
output: pdf_document
---

@@

## Set paths and load data
base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
supplement_figures_dir <- paste0(base_dir, "outputs/figures/supplement/")
tables_dir <- paste0(base_dir, "outputs/tables/")
analysis_objects_dir <- paste0(base_dir, "outputs/objects/")

@@

## Load data objects for plots
objects <- list.files(analysis_objects_dir)
objects <- objects[!str_detect(objects, "null|archive")]
for (obj in objects){
  print(paste0("loading ", obj, "...."))
  load(paste0(analysis_objects_dir, obj))
}

## Load Nirmala's DEG results
df_akula_res <- read_xlsx(path = paste0(base_dir, "data/Akula2021_geneLevel_21kgenes_baseMeanGe5_cqn-Rin-Race-GC_DESeq2-pvalsCorrected_betaPriorT-lfcShrink_092518.xlsx"),
          sheet = "SczVsCtrls") %>% 
    clean_names %>% 
    dplyr::select(ensembl_gene_id, hgnc_symbol, log2fold_change, fold_change, lfc_se, stat, pvalue, padj)
```

@@

```{r TableS1}
write_xlsx(df_covariates, path = paste0(tables_dir, "TableS1.covariate_information.xlsx"))
```

```{r TableS2}
write_xlsx(list("A.GRCCA results" = df_results,
                "B.Covariate (Y) results" = df_y_res,
                "C.Gene (X) results" = df_x_res), 
           path = paste0(tables_dir, "TableS2.GRCCA_results.xlsx")
)
```

@@

```{r TableS4}
write_xlsx(df_de_res, path = paste0(tables_dir, "TableS4.DEGs.xlsx"))
```

@@

```{r TableS6}
write_xlsx(list("A.GRCCA results" = df_results_transcripts,
                "B.Covariate (Y) results" = df_y_res_transcripts,
                "C.Gene (X) results" = df_x_res_transcripts), 
           path = paste0(tables_dir, "TableS6.GRCCA_transcript_results.xlsx")
)
```

```{r TableS7}
write_xlsx(list(
    "A. Module gene assignments" = df_modules_filt %>% left_join(df_ensembl_to_symbol) %>% dplyr::select(module, color, ensembl_gene_id, gene_symbol) %>% as.data.frame,
    "B. Module GO results" = df_mods_go %>% dplyr::filter(pvalue < 0.05 & module != 0) %>% as.data.frame
),
path = paste0(tables_dir, "TableS7.WGCNA_module_results.xlsx")
)
```

```{r TableS8}
write_xlsx(df_lm_dx %>% dplyr::select(-data) %>% dplyr::filter(p_value < 0.05),
           path = paste0(tables_dir, "TableS8.kME_analysis_significant_results.xlsx")
)
```

```{r TableS9}
write_xlsx(list("A.RCCA results" = df_results_rcca,
                "B.Covariate (Y) results" = df_y_res_rcca,
                "C.Gene (X) results" = df_x_res_rcca), 
           path = paste0(tables_dir, "TableS9.RCCA_results.xlsx")
)
```

@@

### B | Correlate qSVs with known covariates

```{r FigS2b.prep}
## Define covariate order based on clustering within each type
f_define_order <- function(covariate_matrix, covariate_type = c("technical", "biological", "toxicological")) {
    if (covariate_type == "technical") {
        subset <- technical_covariates
    } else if (covariate_type == "biological") {
        subset <- biological_covariates
    } else if (covariate_type == "toxicological") {
        subset <- drug_covariates
    } else (stop("Covariate type must be technical, biological, or toxicological"))
    matrix_tmp <- covariate_matrix[subset,]
   order <- matrix_tmp[hclust(dist(matrix_tmp))$order,] %>% rownames
   return(order)
}
covariate_order <- c(
    f_define_order(m_covariates, "biological"),
    f_define_order(m_covariates, "toxicological"),
    f_define_order(m_covariates, "technical")
)
covariate_order

@@

# FIG S3: Drug Multiple Correspondence Analysis results #

@@

```{r FigS3d.plot, fig.width = 3.75, fig.height = 5}
df_figS3d %>% 
    ggplot(aes(x = dimension, y = covariate)) +
    geom_tile(aes(fill = pearsons_r), linewidth = 0.5) +
    geom_text(aes(label = significance), vjust = 0.75, size = 2.5) +
    facet_wrap(vars(type), scales = "free_y", ncol = 1) +
    force_panelsizes(rows = c(length(biological_covariates), length(drug_covariates))) +
    scale_fill_gradientn(colors = rev(brewer.rdbu(100)), limits = c(-1.0, 1.0)) +
    scale_color_manual(values = c("yes" = "black", "no" = "transparent"), guide = "none") +
    guides(fill = guide_colorbar(title = "Pearson's r", title.position = "left", title.hjust = 0.5)) +
    labs(x = "Dimension", y = NULL, 
         title = "D | MCA dimension \ncorrelations with \nknown covariates") +
    theme(
        legend.position = c(-0.80, 1),
        legend.key.height = unit(0.6, "cm"),
        legend.key.width = unit(0.15, "cm"),
        legend.title = element_text(angle = 90),
        axis.text.y = element_text(size = 8.5),
        axis.text.x = element_text(angle = 45, hjust = 1)
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S3D.MCA_dimension_covariate_correlations", .x), width = 3.75, height = 5)
)
```

@@

```{r FigS5b.prep}
## Extract topic descriptions (frequent words)
# FREX weights words by frequency and exclusivity to the topic
df_word_beta <- tidy(topic_model, matrix = "beta") %>%
    group_by(topic) %>%
    arrange(topic, -beta) %>% 
    top_n(n = 10, wt = beta) %>%
    mutate(topic = paste0("topic ", topic)) %>% 
    dplyr::filter(!str_detect(term, "cell")) %>% 
    group_by(topic) %>% 
    mutate(top_word = ifelse(beta == max(beta), "yes", "no"))

@@

# FIG S8: Vector of significant GRCCA genes (FDR_r < 0.05 & z-score > 2) #

@@

# FIG S9: Compare current DEGs to Akula et al 2021 DEGs #

@@

# Akula DEGs vs my DEGs
hypergeometric_overlap <- f_hypergeometric(my_degs, akula_degs, gene_universe = ensembl_gene_ids)
length(hypergeometric_overlap$overlap_genes) # N = 433; p < 0.0001

>>>>

---
title: "Supplemental figures for RLSmith Transl Psych 2025"
author: "Rachel L Smith"
output: pdf_document
---

@@

## Set paths and load data
base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
supplement_figures_dir <- paste0(base_dir, "outputs/figures/supplement/panels/")
tables_dir <- paste0(base_dir, "outputs/tables/")
analysis_objects_dir <- paste0(base_dir, "outputs/objects/")

@@

## Load data objects for plots
objects <- list.files(analysis_objects_dir)
objects <- objects[!str_detect(objects, "null|archive")]
for (obj in objects){
  print(paste0("loading ", obj, "...."))
  load(paste0(analysis_objects_dir, obj))
}
```

@@

```{r TableS1}
write_xlsx(df_covariates %>% 
               mutate(dx = factor(dx, levels = names(dx_colors))) %>% 
               arrange(dx, sex_at_birth, race, age_death) %>% 
               mutate_at(vars(c(bmi, height, weight)), ~ as.numeric(.x)) %>% 
               dplyr::select(-suicide, -sample), 
           path = paste0(tables_dir, "TableS1.covariate_information.xlsx"))
```

```{r TableS2}
write_xlsx(list("A.GRCCA results" = df_results %>% dplyr::select(-set) %>% dplyr::select(varx, everything()),
                "B.Covariate (Y) results" = df_y_res %>% dplyr::rename("structure_correlation" = "pearsons_r"),
                "C.Gene (X) results" = df_x_res %>% dplyr::rename("structure_correlation" = "pearsons_r")
                ), 
           path = paste0(tables_dir, "TableS2.GRCCA_results.xlsx")
)
```

@@

```{r TableS4}
write_xlsx(
    list("TableS4" = df_limma_res %>% arrange(-abs(t))), 
    path = paste0(tables_dir, "TableS4.DEGs.xlsx")
)
```

@@

```{r TableS6}
write_xlsx(list("A.GRCCA results" = df_results_transcripts %>% dplyr::select(-set) %>% dplyr::select(varx, everything()),
                "B.Covariate (Y) results" = df_y_res_transcripts %>% dplyr::rename("structure_correlation" = "pearsons_r"),
                "C. Transcript (X) results" = df_x_res_transcripts %>% dplyr::rename("structure_correlation" = "pearsons_r")), 
           path = paste0(tables_dir, "TableS6.GRCCA_transcript_results.xlsx")
)
```

```{r TableS7}
write_xlsx(list(
    "A. Module gene assignments" = df_modules_filt %>% left_join(df_ensembl_to_symbol) %>% dplyr::select(module, color, ensembl_gene_id, gene_symbol) %>% as.data.frame,
    "B. Module GO results" = df_mods_go %>% dplyr::filter(pvalue < 0.05 & module != 0) %>% dplyr::select(-qvalue) %>% as.data.frame
),
path = paste0(tables_dir, "TableS7.WGCNA_module_results.xlsx")
)
```

```{r TableS8}
write_xlsx(df_kme %>% 
               dplyr::filter(p_value < 0.05 & pc == "PC1") %>% 
               dplyr::select(-c(data, pc, proportion_of_variance, estimate, std_error)),
           path = paste0(tables_dir, "TableS8.kME_analysis_significant_results.xlsx")
)
```

```{r TableS9}
write_xlsx(list("A.RCCA results" = df_results_rcca %>% dplyr::select(-set) %>% dplyr::select(varx, everything()),
                "B.Covariate (Y) results" = df_y_res_rcca %>% dplyr::rename("structure_correlation" = "pearsons_r"),
                "C.Gene (X) results" = df_x_res_rcca %>% dplyr::rename("structure_correlation" = "pearsons_r")), 
           path = paste0(tables_dir, "TableS9.RCCA_results.xlsx")
)
```

@@

### B | Correlate qSVs with known covariates
```{r FigS2b.prep}
## Define covariate order based on clustering within each type
f_define_order <- function(covariate_matrix, covariate_type = c("technical", "biological", "toxicological")) {
    if (covariate_type == "technical") {
        subset <- technical_covariates
    } else if (covariate_type == "biological") {
        subset <- biological_covariates
    } else if (covariate_type == "toxicological") {
        subset <- drug_covariates
    } else (stop("Covariate type must be technical, biological, or toxicological"))
    matrix_tmp <- covariate_matrix[,subset]
   order <- matrix_tmp[hclust(dist(matrix_tmp))$order,] %>% colnames
   return(order)
}
m_covariates <- df_covariates_numeric %>% dplyr::select(technical_covariates, biological_covariates, drug_covariates) %>%  as.matrix()
covariate_order <- c(
    f_define_order(m_covariates, "biological"),
    f_define_order(m_covariates, "toxicological"),
    f_define_order(m_covariates, "technical")
)
covariate_order

@@

# FIG S3: Toxicology Multiple Correspondence Analysis results #

@@

```{r FigS3d.plot, fig.width = 3.75, fig.height = 5}
df_figS3d %>% filter(type != "Diagnostic group") %>%

    ggplot(aes(x = dimension, y = covariate)) +
    geom_tile(aes(fill = pearsons_r), linewidth = 0.5) +
    geom_text(aes(label = significance), vjust = 0.75, size = 2.5) +
    facet_wrap(vars(type), scales = "free_y", ncol = 1) +
    force_panelsizes(rows = c(length(biological_covariates), length(drug_covariates))) +
    scale_fill_gradientn(colors = rev(brewer.rdbu(100)), limits = c(-1.0, 1.0)) +
    scale_color_manual(values = c("yes" = "black", "no" = "transparent"), guide = "none") +
    guides(fill = guide_colorbar(title = "Pearson's r", title.position = "left", title.hjust = 0.5)) +
    labs(x = "Dimension", y = NULL, 
         title = "D | MCA dimension \ncorrelations with \nknown covariates") +
    theme(
        legend.position = c(-0.80, 1),
        legend.key.height = unit(0.6, "cm"),
        legend.key.width = unit(0.15, "cm"),
        legend.title = element_text(angle = 90),
        axis.text.y = element_text(size = 8.5),
        axis.text.x = element_text(angle = 45, hjust = 1)
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S3D.MCA_dimension_covariate_correlations", .x), width = 3.75, height = 5)
)
```

@@

```{r FigS5b.prep}
n_topics <- topic_model$settings$dim$K
## Extract topic descriptions (frequent words)
# FREX weights words by frequency and exclusivity to the topic
df_word_beta <- tidy(topic_model, matrix = "beta") %>%
    group_by(topic) %>%
    arrange(topic, -beta) %>% 
    top_n(n = 10, wt = beta) %>%
    mutate(topic = paste0("topic ", topic)) %>% 
    dplyr::filter(!str_detect(term, "cell")) %>% 
    group_by(topic) %>% 
    mutate(top_word = ifelse(beta == max(beta), "yes", "no"))

@@

# FIG S8: Vector of significant GRCCA genes (FDR_r < 0.05 & Z-score > 2) #

@@

# FIG S9: Compare across WGCNA module solutions

### A | Model significance
```{r FigS9a, fig.width = 6.5, fig.height = 2}
df_model_res_wgcna_groups %>% 
    group_by(mod_set) %>% 
    mutate(best = ifelse(pval == min(pval), 1, 0)) %>% 
    mutate(best = ifelse(set == 5, 0, best)) %>% 
    mutate(best = factor(best)) %>%
    mutate(mod_set = str_replace(mod_set, "_", ", ")) %>%

    ggplot(aes(x = factor(varx), y = -log10(pval))) +
    geom_hline(yintercept = -log10(0.005), color = "grey", linewidth = 0.5) +
    geom_segment(aes(y = 0, yend = -log10(pval), color = mod_set, linetype = mod_set),
                 position = position_dodge(width = 0.6)
    ) +
    geom_point(aes(fill = mod_set, size = correl, shape = best), 
               position = position_dodge(width = 0.6)
    ) +
    scale_size_continuous(range = c(1, 4)) +
    scale_shape_manual(values = c("0" = 21, "1" = 23), guide = "none") +
    scale_color_manual(values = c("actual (minSize40, cutHeight0.98)" = "#F8766D", rep("gray", 3))) +
    scale_fill_manual(values = c("actual (minSize40, cutHeight0.98)" = "#F8766D", rep("gray", 3))) +
    scale_linetype_manual(values = c("actual (minSize40, cutHeight0.98)" = "solid", 
                                     "minSize30, cutHeight0.95" = "dashed", 
                                     "minSize30, cutHeight0.98" = "dotted", 
                                     "minSize40, cutHeight0.95" = "twodash")
    ) +
    guides(fill = guide_legend(title = NULL), color = guide_legend(title = NULL),
           size = guide_legend(title = "X-Y cor"), linetype = guide_legend(title = NULL)) +
    coord_cartesian(clip = "off") +
    labs(x = "X variance explained", y = "-log10(model P value)",
         title = "A | Model significance across WGCNA solutions") +
    theme(
        legend.key.height = unit(0.2, "cm"),
        legend.title = element_text(size = 8),
        legend.position = c(0.025, 0.68),
        legend.spacing = unit(0.2, "lines")
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S9A.wgcna_groups_model_significance", .x), width = 6.5, height = 2)
)
```

### B | X and Y results correlations
```{r FigS9b, fig.width = 4.0, fig.height = 3.75}
# Plot
plot_y_func <- function(data, mapping, ...) {
    ggplot(data = data, mapping = mapping) +
        geom_point(alpha = 0.5) +
        #geom_smooth(method = "lm", color = "maroon", se = FALSE, linewdith = 0.25) +
        stat_cor(aes(label = ..r.label..), size = 3, color = "maroon")
}
plot_x_func <- function(data, mapping, second_df, ...) {
    ggplot(data = second_df, mapping = mapping) +  
        geom_point(alpha = 0.5, size = 0.5) +
        stat_cor(aes(label = ..r.label..), size = 3, label.x.npc = 0.4, label.y.npc = 0.1, color = "maroon")
}

ggpairs(df_y_res_wgcna_groups %>% 
            dplyr::rename_all(~str_replace(.x, "_", ",\n")) %>% 
            dplyr::rename_at(4, ~"actual"),
        lower = list(continuous = plot_y_func),  # Scatter + best fit line
        upper = list(continuous = wrap(plot_x_func, 
                                       second_df = df_x_res_wgcna_groups %>% 
                                           dplyr::rename_all(~str_replace(.x, "_", ",\n")) %>% 
                                           dplyr::rename_at(4, ~"actual")
                                       )
        ),  # Correlation values in upper triangle
        diag = list(continuous = wrap("blank"))  # Remove density plot from diagonal
) + 
    theme_bw() +
    labs(title = "B | Model feature correlations") +
    theme(plot.title = element_text(face = "bold", size = 11),
          strip.text = element_text(size = 8.25),
          strip.clip = "off",
          strip.background = element_blank(),
          panel.spacing = unit(0.2, "lines")
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S9B.wgcna_groups_x_y_cor", .x), width = 4.0, height = 3.75)
)
```

### C | Enrichment for risk genes
```{r FigS9c, fig.height = 3.75, fig.width = 2.4}
df_benchmarking_res_wgcna_groups %>%
    mutate(significant = ifelse(padj < 0.05, "yes", "no")) %>%
    mutate(class = ifelse(str_detect(benchmark_list, "SCZ"), 
                          "SCZ",
                          benchmark_list),
           benchmark_list = ifelse(str_detect(benchmark_list, "SCZ"),
                                   str_remove(benchmark_list, "SCZ \\(") %>% str_remove("\\)") %>% str_replace(", ", ",\n"), 
                                   ""
           ),
           mod_set = str_replace(mod_set, "_", ",\n")
    ) %>%

    ggplot(aes(x = benchmark_list, y = mod_set)) +
    geom_point(aes(fill = NES, size = -log10(padj), color = significant), 
               stroke = 1, shape = 21) +

    facet_wrap(vars(class), ncol = 1, scales = "free_y") +
    force_panelsizes(rows = c(1,1,1,3)) +
    scale_size_continuous(range = c(0.5, 7)) +
    scale_color_manual(values = c("yes" = "black", "no" = "transparent"), guide = "none") +
    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-1.61, 1.61)) +
    guides(fill = guide_colorbar(title = "NES", title.position = "left", title.hjust = 0.5), 
           size = guide_legend(title = "-log10(FDR)", title.position = "left", title.hjust = 0.5)) +
    coord_flip(clip = "off") +
    labs(x = NULL, y = NULL,
         title = "C | Risk gene enrichments \nacross WGCNA solutions") +
    theme(
        legend.key.height = unit(0.5, "cm"),
        legend.key.width = unit(0.25, "cm"),
        legend.title = element_text(size = 8, angle = 90, margin = margin(r = 0)),
        legend.text = element_text(margin = margin(r = -1)),
        legend.spacing = unit(0.2, "lines"),
        legend.position = c(-0.32, 0.75),
        panel.spacing = unit(0.05, "lines"),
        strip.text = element_text(size = 9),
        axis.text.x = element_text(size = 9, angle = 45, hjust = 0.9),
        plot.margin = margin(b = -10)
        #legend.margin = margin(l = -10)
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S9C.wgcna_groups_benchmarking", .x), width = 2.4, height = 3.75)
)
```

# FIG S10: Compare across random module solutions

### A | Model significance
```{r FigS10a, fig.width = 6.5, fig.height = 2}
seeds <- df_model_res_random_groups %>% filter(seed != "actual") %>% pull(seed) %>% unique
df_model_res_random_groups %>% 
    group_by(seed) %>% 
    mutate(best = ifelse(pval == min(pval), 1, 0)) %>%
    mutate(best = ifelse(seed %in% c("seed1856", "seed3144", "seed3578", "seed9883") & set == 4, 0, best)) %>%
    mutate(best = factor(best)) %>%
    arrange(desc(seed)) %>%

    ggplot(aes(x = factor(varx), y = -log10(pval))) +
    geom_hline(yintercept = -log10(0.005), color = "grey", linewidth = 0.5) +
    geom_segment(aes(y = 0, yend = -log10(pval), color = seed),
                 position = position_dodge(width = 0.80)
    ) +
    geom_point(aes(fill = seed, size = correl, shape = best), 
               position = position_dodge(width = 0.80)
    ) +
    scale_color_manual(values = c("actual" = "#F8766D", rep("gray", 10))) +
    scale_fill_manual(values = c("actual" = "#F8766D", 
                                 setNames(rep(adjustcolor("gray", alpha.f = 0.5), 10), seeds)),
                      guide = "none"
    ) +
    scale_size_continuous(range = c(1, 4)) +
    scale_shape_manual(values = c("0" = 21, "1" = 23), guide = "none") +
    guides(color = guide_legend(title = NULL), size = guide_legend(title = "X-Y cor")) +
    coord_cartesian(clip = "off") +
    labs(x = "X variance explained", y = "-log10(model P value)",
         title = "A | Model significance across random groupings") +
    theme(
        legend.key.height = unit(0.2, "cm"),
        legend.title = element_text(size = 8),
        legend.position = c(0.025, 0.60),
        legend.spacing = unit(0.5, "lines")
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S10A.random_groups_model_significance", .x), width = 6.5, height = 2)
)
```

### B | X and Y results correlations
```{r FigS10b, fig.width = 6.5, fig.height = 2}
df_y_res_random_groups %>% 
    pivot_longer(contains("seed"), names_to = "seed", values_to = "random") %>%
    mutate(type = "Covariates \n(ry)") %>%

    bind_rows(
        df_x_res_random_groups %>% 
            pivot_longer(contains("seed"), names_to = "seed", values_to = "random") %>% 
            mutate(type = "Genes \n(rx)")
    ) %>%

    ggplot(aes(x = random, y = actual)) +
    geom_point(alpha = 0.5, size = 0.5) +
    stat_cor(aes(label = ..r.label..), label.y = 0.8, color = "maroon", size = 2.75) +
    facet_grid(type ~ seed) +
    scale_x_continuous(breaks = c(-0.5, 0.5)) +
    labs(x = "Random grouping structure correlations", y = "Actual structure correlations",
         title = "B | Model feature correlations") +
    coord_cartesian(clip = "off") +
    theme_bw() +
    theme(
        strip.text.x = element_blank(),
        strip.text.y = element_text(size = 8.5),
        strip.background = element_blank(),
        strip.clip = "off",
        plot.title = element_text(face = "bold", size = 11),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 8),
        panel.border = element_rect(linewidth = 0.30),
        plot.margin = margin(r = -5)
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S10B.random_groups_x_y_cor", .x), width = 6.5, height = 2)
)
```

### C | Enrichment for risk genes
```{r FigS10c, fig.height = 2, fig.width = 6.5}
df_benchmarking_res_random_groups %>%
    mutate(significant = ifelse(padj < 0.05, "yes", "no")) %>%
    mutate(class = ifelse(str_detect(benchmark_list, "SCZ"), 
                          "SCZ",
                          benchmark_list),
           benchmark_list = ifelse(str_detect(benchmark_list, "SCZ"),
                                   str_remove(benchmark_list, "SCZ \\(") %>% str_remove("\\)") %>% str_replace(", ", ",\n"), 
                                   ""
           )
    ) %>%

    ggplot(aes(x = benchmark_list, y = seed)) +
    geom_point(aes(fill = NES, size = -log10(padj), color = significant), 
               stroke = 1, shape = 21) +

    facet_wrap(vars(class), nrow = 1, scales = "free_x") +
    force_panelsizes(cols = c(1,1,1,3)) +
    scale_size_continuous(range = c(0.1, 5)) +
    scale_color_manual(values = c("yes" = "black", "no" = "transparent"), guide = "none") +
    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-1.61, 1.61)) +
    guides(fill = guide_colorbar(title = "NES", title.position = "left", title.hjust = 0.5), 
           size = guide_legend(title = "-log10(FDR)", title.position = "left", title.hjust = 0.5)) +
    coord_cartesian(clip = "off") +
    labs(x = NULL, y = NULL,
         title = "C | Risk gene enrichments across random groupings") +
    theme(
        legend.key.height = unit(0.5, "cm"),
        legend.key.width = unit(0.25, "cm"),
        legend.title = element_text(size = 8, angle = 90),
        legend.spacing = unit(0.2, "lines"),
        legend.margin = margin(l = -10)
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S10C.random_groups_benchmarking", .x), width = 6.5, height = 2)
)
```

# FIG S11: Compare GRCCA and RCCA directly #

### A | Correlate GRCCA and RCCA covariate (y) structure correlations
```{r FigS11a.plot}
df_y_res_all %>% 
    pivot_wider(id_cols = covariate, names_from = analysis, values_from = pearsons_r) %>% 
    ggplot(aes(x = rcca, y = grcca)) +
    geom_point(aes(fill = rcca + grcca), shape = 21, size = 2.5) +
    geom_vline(aes(xintercept = 0), color = "gray") +
    geom_hline(aes(yintercept = 0), color = "gray") +
    geom_abline(lty = 2) +
    geom_text_repel(aes(label = covariate), size = 2.5, max.overlaps = 20, box.padding = 1.25,
                     nudge_y = 0.3, force = 150) +
    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-1.84, 1.84), guide = "none") +
    labs(title = "A | Covariate structure correlations",
         x = "RCCA", y = "GRCCA")
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S11A.covariate_structCor_correlation", .x), width = fig_width, height = fig_height)
)
```

### B | Correlate GRCCA and RCCA gene (x) structure correlations
```{r FigS11b.plot}
df_x_res_all %>% 
    pivot_wider(id_cols = ensembl_gene_id, names_from = analysis, values_from = pearsons_r) %>% 
    ggplot(aes(x = rcca, y = grcca)) +
    geom_point(aes(color = rcca + grcca), size = 1) +
    geom_vline(aes(xintercept = 0), color = "gray") +
    geom_hline(aes(yintercept = 0), color = "gray") +
    geom_abline(lty = 2) +
    stat_cor(label.sep = "\n") +
    #geom_label_repel(aes(label = gene_symbol), size = 3, max.overlaps = 50, box.padding = 1.5) +
    scale_color_gradientn(colors = gene_weight_color_scale, limits = c(-1.14, 1.14), guide = "none") +
    labs(title = "B | Gene structure correlations",
         x = "RCCA", y = "GRCCA")
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S11B.gene_structCor_correlation", .x), width = fig_width, height = fig_height)
)
```

### C | Venn diagram of significant gene overlap (made in PowerPoint/Inkscape)
```{r FigS11c.prep}
## Venn diagram (overlap with Nirmala's results)
sig_grcca_genes <- df_x_res %>%
    dplyr::filter(p_adj < 0.05 & abs(z_score) > 2) %>%
    pull(ensembl_gene_id)
sig_rcca_genes <- df_x_res_rcca %>% 
    dplyr::filter(p_adj < 0.05 & abs(z_score) > 2) %>% 
    pull(ensembl_gene_id)
length(sig_grcca_genes) # N = 1211
length(sig_rcca_genes) # N = 2026
hypergeometric_overlap <- f_hypergeometric(sig_grcca_genes, sig_rcca_genes, gene_universe = ensembl_gene_ids)
length(hypergeometric_overlap$overlap_genes) # N = 855; P < 0.0001
```

### D | RCCA results risk gene enrichment
```{r FigS11d.prep}
df_fig13d <- df_rcca_fgsea_benchmark %>% 
    mutate(class = ifelse(str_detect(benchmark_list, "SCZ"), "SCZ", "other") %>% factor(levels = c("SCZ", "other"))) %>% 
    mutate(benchmark_list = str_remove_all(benchmark_list, "SCZ|\\)|\\(") %>% 
               str_trim %>% 
               str_replace(", ", ",\n")
    ) %>% 
    mutate(label_color = ifelse(padj < 0.05, "white", "black")) %>% 
    mutate(box_color = ifelse(padj < 0.05, "black", "transparent")) %>% 
    mutate(label = round(NES, 2) %>% format(digits = 3))
    #mutate(sig = ifelse(padj < 0.05, scientific(padj, digits = 3), round(padj, 2) %>% format(digits = 3)))
   # mutate(sig = case_when(padj < 0.001 ~ "***", padj < 0.01 ~ "**", padj < 0.05 ~ "*", TRUE ~ ""))
```

```{r FigS11d.plot}
df_fig13d %>% 
    ggplot(aes(x = benchmark_list, y = 1)) +
    geom_tile(aes(fill = NES, color = I(box_color)), width = 0.95, height = 0.95, linewidth = 0.75) +
    geom_text(aes(label = label, color = I(label_color)), size = 3) +
    facet_wrap(vars(class), scales = "free", nrow = 2) +
    scale_fill_gradient(low = "white", high = "#67001F", limits = c(0.838, 1.58)) +
    guides(fill = guide_colorbar(title.hjust = 0.5, title.position = "left")) +
    labs(x = NULL, y = NULL,
         title = "D | Risk gene enrichment") +
    theme(axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          axis.line = element_blank(),
          axis.text.x = element_text(margin = margin(t = -3)),
          strip.background = element_rect(fill = "white", color = "black"),
          legend.margin = margin(r = -15),
          legend.position = "left",
          legend.justification = "top",
          legend.title = element_text(angle = 90),
          legend.key.height = unit(0.75, "cm"),
          legend.key.width = unit(0.15, "cm")
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S11D.Risk_gene_enrichment", .x), width = fig_width, height = fig_height)
)
```

# FIG S12: Compare limme DEG results with Akula et al 2021 #

### A | Venn diagram of significant DEG overlap
```{r FigS12a.prep}
## Venn diagram (overlap with Nirmala's results)
my_degs <- df_limma_res %>% dplyr::filter(p_value < 0.05) %>% pull(ensembl_gene_id)
length(my_degs) # N = 1389

@@

# Akula DEGs vs my DEGs
hypergeometric_overlap <- f_hypergeometric(my_degs, akula_degs, gene_universe = ensembl_gene_ids)
length(hypergeometric_overlap$overlap_genes) # N = 385; P = 1.9*10^-145; OR = 7.17

# GRCCA vs my DEGs
hypergeometric_overlap <- f_hypergeometric(my_degs, sig_grcca_genes, gene_universe = ensembl_gene_ids)
length(hypergeometric_overlap$overlap_genes) # N = 203; P = 1.5*10^-29; OR = 2.77

# GRCCA vs Akula DEGs
hypergeometric_overlap <- f_hypergeometric(akula_degs, sig_grcca_genes, gene_universe = ensembl_gene_ids)
length(hypergeometric_overlap$overlap_genes) # N = 285; P = 2.3*10^-85; OR = 5.19

# Combine the data into a named list for ggvenn
FigS12a_data <- list(
  `Current study DEGs` = my_degs,
  `Akula et al 2021 DEGs` = akula_degs,
  `Significant GRCCA genes` = sig_grcca_genes
)
```

```{r FigS12a.plot}
ggvenn(
    FigS12a_data, 
    fill_color = c("#E0E0E0", "#E0E0E0", "#E0E0E0"),
    stroke_size = 0.75,
    text_size = 3,
    set_name_size = 4,
    show_percentage = FALSE
) +
    coord_fixed(clip = "off") +
    theme_void() +
    theme(
        text = element_text(size = 8)
    )
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S12A.DEG_GRCCA_venn_diagram", .x), width = fig_width, height = fig_height)
)
```

### B | Correlation plot of L2FC
```{r FigS12b.stats}

### PERMUTE L2FC CORRELATION ###
n_perm <- 10000
df_limma_akula_cor_perm <- tibble()
set.seed(0406)
for (i in 1:n_perm) {

    if (i %% 100 == 0) {print(paste0("Running permutation ", i))}

    # Permute LFC within module
    df_tmp <- df_de_res_all %>% 
        left_join(df_modules_filt) %>%
        filter(!is.na(Akula) & metric == "l2fc") %>% 
        group_by(module) %>%
        mutate(permuted_l2fc = sample(Limma, replace = FALSE))

    # Calculate correlation between permuted data and Nirmala's LFC
    test <- cor.test(df_tmp$permuted_l2fc, df_tmp$Akula)
    correlation <- test$estimate
    p_value <- test$p.value

    # Append to tibble
    df_limma_akula_cor_perm <- df_limma_akula_cor_perm %>% 
        bind_rows(
            tibble(
                perm = i,
                correlation = correlation,
                p_value = p_value
            )
        )

}

## Calculate permutation p-values
de_limma_akula_perm_p <- df_limma_akula_cor_perm %>% 
    mutate(obs_corr = cor.test(
        df_de_res_all %>% filter(!is.na(Akula) & metric == "l2fc") %>% pull(Limma), 
        df_de_res_all %>% filter(!is.na(Akula) & metric == "l2fc") %>% pull(Akula), 
        )$estimate
    ) %>%
    summarise(
        perm_p = (sum(abs(correlation) >= abs(obs_corr)) + 1) / (n_perm + 1)
    )
#Pperm=1*10^-4

### PERMUTE T-STAT CORRELATION ###
n_perm <- 10000
df_limma_akula_cor_perm <- tibble()
set.seed(0406)
for (i in 1:n_perm) {

    if (i %% 100 == 0) {print(paste0("Running permutation ", i))}

    # Permute LFC within module
    df_tmp <- df_de_res_all %>% 
        left_join(df_modules_filt) %>%
        filter(!is.na(Akula) & metric == "t_stat") %>% 
        group_by(module) %>%
        mutate(permuted_t = sample(Limma, replace = FALSE))

    # Calculate correlation between permuted data and Nirmala's LFC
    test <- cor.test(df_tmp$permuted_t, df_tmp$Akula)
    correlation <- test$estimate
    p_value <- test$p.value

    # Append to tibble
    df_limma_akula_cor_perm <- df_limma_akula_cor_perm %>% 
        bind_rows(
            tibble(
                perm = i,
                correlation = correlation,
                p_value = p_value
            )
        )

}

## Calculate permutation p-values
de_limma_akula_perm_p <- df_limma_akula_cor_perm %>% 
    mutate(obs_corr = cor.test(
        df_de_res_all %>% filter(!is.na(Akula) & metric == "t_stat") %>% pull(Limma), 
        df_de_res_all %>% filter(!is.na(Akula) & metric == "t_stat") %>% pull(Akula), 
        )$estimate
    ) %>%
    summarise(
        perm_p = (sum(abs(correlation) >= abs(obs_corr)) + 1) / (n_perm + 1)
    )
#Pperm=1*10^-4

```{r FigS12b.prep}
## Create df for plot
df_figS12b <- df_de_res_all %>%
